---
title: "README"
author: "Mark Ravina"
date: "`r Sys.Date()`"
# output: html_document
output: md_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# sfhelper

License: MIT

For more information please contact mark.ravina@austin.utexas.edu

## Installation

To install from GitHub use the devtools package

```{r eval=FALSE}
devtools::install_github("histmr/sfhelper")
```
## Functions

### geolocate()

The geolocate() function takes a data frame with a column of place names (toponyms) and a column of regions, specified by their two-digits ISO codes. The defualt colums names are "place" and "iso". The function uses the API of the World Historical Gazeteer (https://whgazetteer.org/) to generate a four column data frame with the toponym, ISO code, longitude, and latitude. Many toponyms will retrun multple hits. You will probably want to import a data frame, but here's an example of a data frame create within R

```{r}
library(sfhelper)
df <- data.frame("place"=c("Tokyo","Edo","Prague","Prague"),"iso"=c("JP","JP","CZ",""))
geolocate(df,place,iso)
```

### st_transform_repair

Changing the meridian when re-projecting sf object often creates broken polygons and other unwanted artifacts. This function repairs most common errors. For example, using the map  from **rnaturalearth**, the **st_transform_repair()** function corrects for the new meridian.

```{r, warning=FALSE,message=FALSE}
library("rnaturalearth")
library("tidyverse")
library("sf")

world.sf <- ne_countries(scale = "medium", returnclass = "sf")

new_crs <- "+proj=moll +lat_1=-10 +lon_0=-70"

new.sf <- st_transform(x = world.sf, crs = new_crs)
ggplot() + geom_sf(data=new.sf)

new.sf <- st_transform_repair(x = world.sf, crs= new_crs)
ggplot() + geom_sf(data=new.sf)
```

In an orthographic projection, the US disappears if we re-project to "+proj=ortho +lat_0=40 +lon_0=-12".

```{r, warning=FALSE,message=FALSE}
new_crs <- "+proj=ortho +lat_0=40 +lon_0=-12"
ggplot() + geom_sf(data= world.sf %>% st_transform(new_crs))
```

The **st_transform_repair()** fixes the broken polygons

```{r, warning=FALSE,message=FALSE}
ggplot() + geom_sf(data=world.sf %>% st_transform_repair(new_crs))
```

### st_transform_outline

As in the examples above, re-projected maps sometimes lack an "edge of the globe" graticule line. The **st_transform_outline()** function generates that line as an **sf** object. Using with **fill** in **ggplot()**, the new object can be used to represent oceans. At present it does NOT work for orthographic projections with changes in **+lat_0=0**

```{r, warning=FALSE,message=FALSE}
new_crs <- "+proj=ortho +lat_0=0 +lon_0=-12"
ggplot() + 
  geom_sf(data=st_transform_outline(new_crs), fill="lightblue") +
  geom_sf(data=world.sf %>% st_transform_repair(new_crs))


new_crs <- "+proj=eck4 +lon_0=-70"
ggplot() + 
    geom_sf(data=st_transform_outline(new_crs), fill="lightblue") +
    geom_sf(data=world.sf %>% st_transform_repair(new_crs))
```

These transforms make the graticule disappear, but we can restore it with **st_graticule()**

```{r, warning=FALSE,message=FALSE}
new_crs <- "+proj=ortho +lon_0=-20"
new.sf <- world.sf %>% st_transform_repair(new_crs)
new_graticule.sf <- st_graticule(new.sf) %>% st_transform(new_crs)
ggplot() + geom_sf(data=st_transform_outline(crs=new_crs), fill="lightblue") +
  geom_sf(data=new.sf) +
  geom_sf(data=new_graticule.sf, color="white", linewidth=0.3, alpha=0.7)
```

