---
title: "README"
author: "Mark Ravina"
date: "`r Sys.Date()`"
output: html_document
# output: md_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# sfhelper

License: MIT

For more information please contact mark.ravina@austin.utexas.edu

## Installation

To install from GitHub use the devtools package

```{r eval=FALSE}
devtools::install_github("histmr/sfhelper")
```
## Functions

### geolocate()

The geolocate() function takes a data frame with a column of place names (toponyms) and a column of regions, specified by their two-digits ISO codes. The defualt colums names are "place" and "iso". The function uses the API of the World Historical Gazeteer (https://whgazetteer.org/) to generate a four column data frame with the toponym, ISO code, longitude, and latitude. Many toponyms will retrun multple hits. You will probably want to import a data frame, but here's an example of a data frame create within R

```{r}
library(sfhelper)
df <- data.frame("place"=c("Tokyo","Edo","Prague","Prague"),"iso"=c("JP","JP","CZ",""))
geolocate(df,place,iso)
```

### st_transform_repair

Changing the meridian when re-projecting sf object often creates broken polygons and other unwanted artifacts. This function repairs most common errors. For example, using the map  from **rnaturalearth**, the **st_transform_repair()** function corrects for the new meridian.

```{r, warning=FALSE,message=FALSE}
library(rnaturalearth)
library(tidyverse)
library(sf)
library(cowplot)

world.sf <- ne_countries(scale = "medium", returnclass = "sf")

new_crs <- "+proj=moll +lat_1=-10 +lon_0=-70"

new.sf <- st_transform(x = world.sf, crs = new_crs)
ggplot() + geom_sf(data=new.sf)

new.sf <- st_transform_repair(x = world.sf, crs= new_crs)
ggplot() + geom_sf(data=new.sf)
```

In an orthographic projection, the US disappears if we re-project to "+proj=ortho +lat_0=40 +lon_0=-12".

```{r, warning=FALSE,message=FALSE}
new_crs <- "+proj=ortho +lat_0=40 +lon_0=-12"
ggplot() + geom_sf(data= world.sf %>% st_transform(new_crs))
```

The **st_transform_repair()** fixes the broken polygons

```{r, warning=FALSE,message=FALSE}
ggplot() + geom_sf(data=world.sf %>% st_transform_repair(new_crs))
```

### st_transform_outline

As in the examples above, re-projected maps sometimes lack an "edge of the globe" graticule line. The **st_transform_outline()** function generates that line as an **sf** object. Using with **fill** in **ggplot()**, the new object can be used to represent oceans. At present it does NOT work for orthographic projections with changes in **+lat_0=0**

```{r, warning=FALSE,message=FALSE}
new_crs <- "+proj=ortho +lat_0=0 +lon_0=-12"
ggplot() + 
  geom_sf(data=st_transform_outline(new_crs), fill="lightblue") +
  geom_sf(data=world.sf %>% st_transform_repair(new_crs))


new_crs <- "+proj=eck4 +lon_0=-70"
ggplot() + 
    geom_sf(data=st_transform_outline(new_crs), fill="lightblue") +
    geom_sf(data=world.sf %>% st_transform_repair(new_crs))
```

These transforms make the graticule disappear, but we can restore it with **st_graticule()**

```{r, warning=FALSE,message=FALSE}
new_crs <- "+proj=ortho +lon_0=-20"
new.sf <- world.sf %>% st_transform_repair(new_crs)
new_graticule.sf <- st_graticule(new.sf) %>% st_transform(new_crs)
ggplot() + geom_sf(data=st_transform_outline(crs=new_crs), fill="lightblue") +
  geom_sf(data=new.sf) +
  geom_sf(data=new_graticule.sf, color="white", linewidth=0.3, alpha=0.7)
```

### st_equal_grid()

This function was inspired by Daniel Immerwahr's *How to Hide an Empire* (2019), in which he combines non-contiguous US states and territories, at the same scale. This  approach is useful for visualizing the relative sizes of geographically disparate state, countries, territories, etc. Following Immerwahr, let's show the continental US, Alaska, Hawaii, and the Philippines at the same scale. The **st_equal_grid()** function takes a list of **sf** objects, all with projected coordinates in the same units. You can (likely will want to) use different projections for different places, but they must all be in the same units

```{r}
library(spData)
logo.sf <- us_states %>% 
  st_transform("+proj=lcc +lon_0=-90 +lat_1=33 +lat_2=45") %>% 
  st_union()## continental US
alaska.sf <- alaska %>% 
  st_transform("+proj=aea +lat_0=50 +lon_0=-154 +lat_1=55 +lat_2=65 +x_0=0 +y_0=0 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs +type=crs") %>% st_union()
hawaii.sf <- hawaii %>% st_transform("+proj=utm +zone=32") %>% st_union()

world_ne.sf <- st_as_sf(ne_countries(scale="medium"))
philippines.sf <- world_ne.sf %>% filter(sov_a3=="PHL") %>% st_union() %>% st_transform(3121) # https://epsg.io/3121

places <- list(philippines.sf,logo.sf,alaska.sf,hawaii.sf)
plot_list  <- st_equal_grid(places)
```

To visualize the plots in a grid, use the **cowplot** package

```{r}
cowplot::plot_grid(plotlist = plot_list, ncol = 4)
```

```{r}
cowplot::plot_grid(plotlist = plot_list, ncol = 2)
```

Using a unprojected **sf** object will generate strange results

```{r}
bad_hawaii.sf <- hawaii.sf %>% st_transform((4326))
places <- list(philippines.sf,logo.sf,alaska.sf,bad_hawaii.sf)
plot_list  <- st_equal_grid(places)
```
